Bayesian exponential random graph modeling of KWL Burial network

### 2020-09-20 analysis/paper/paper.Rmd
minor updates for method
edit the main text

Using the R software, we use Bayesian SNA to study burials from Kiwulan, an Iron Age site in northeast Taiwan. We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group. We test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions. We ask the questions: 1. Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities? 2. Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an indigenous society can be detected by burials networks. 3. How the relations among burials may affect the quantity of burial goods? We expect the network to present a higher degree of clustering (e.g. the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period. This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on indigenous groups. A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.

@@

@@

Markov chain Monte Carlo (MCMC) is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network. Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017]. Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density.

@@

The pre-European network shows three subgroups that possessing the same types of burial goods: medium amount of carnelian bead on the right, medium amount of gold-foil beads on the left, and porcelains in the middle. Although burials with different amount of burial goods can be identified, there is no clear clusters demonstrating the sharing of prestige goods only for burials with rich goods.

>>>>

Using the R software, we apply Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan. We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group. We test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions. We ask the questions: 1. Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities? 2. Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an indigenous society can be detected by burials networks. 3. How the relations among burials may affect the quantity of burial goods? We expect the network to present a higher degree of clustering (e.g. the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period. This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on indigenous groups. A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.

@@

## Bayesian modelling

@@

Markov chain Monte Carlo (MCMC) is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network. Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].

@@

The pre-European network shows three subgroups that possessing the same types of burial goods: medium amount of carnelian bead on the right, medium amount of gold-foil beads on the left, and porcelains in the middle. Although burials with different amount of burial goods can be identified, there is no clear clusters demonstrating the sharing of prestige goods only for burials with rich goods. The post-European network also presented subgroups with more edges connecting burials.
